---
title: "Assignment 5"
author: "Nicholas Irons [njirons@uw.edu]"
date: '`r format(Sys.time(), "%Y-%m-%d")`'
header-includes:
- \usepackage{float} #use the 'float' package
- \floatplacement{figure}{H} #make every figure with caption = h
output: 
    html_document:
        number_sections: true      
        self_contained: true
        toc: true
        toc_float:
            collapsed: true
            smooth_scroll: false
    pdf_document:
        number_sections: true
        toc: true
        fig_caption: yes
        keep_tex: yes
    always_allow_html: true
urlcolor: blue  
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ISOcodes)
library(HMDHFDplus)
library(tidyverse)
library(keyring)
library(ggpubr)

library(captioner)
table_nums <- captioner(prefix = "Table")
figure_nums <- captioner(prefix = "Figure")
```

# HFD exploration

Get country codes and define functions to pull data from HFD:

```{r}
# HFD country codes
hfdcodes <- getHFDcountries() %>% tibble(ccode = .)

# ISO country codes
isocodes <- ISO_3166_1 %>% tibble() %>% select(ccode = Alpha_3, Name)

# join ISO codes with country names
hfdcodes <- hfdcodes %>% left_join(isocodes, by = "ccode")

# update missing country names 
hfdcodes <- hfdcodes %>% 
    mutate(Name = 
        case_when(ccode == "FRATNP" ~  "France",
                  ccode == "DEUTNP" ~  "Germany",
                  ccode == "DEUTE" ~   "East Germany",
                  ccode == "DEUTW" ~   "West Germany",
                  ccode == "GBR_NP" ~  "United Kingdom", 
                  ccode == "GBRTENW" ~ "England and Wales",
                  ccode == "GBR_SCO" ~ "Scotland",
                  ccode == "GBR_NIR" ~ "Northern Ireland",
                  TRUE ~ Name)
    )

# a function to read HFD for one country and one item
read_hfd_country <- function(CNTRY, item) {
  HMDHFDplus::readHFDweb(
    # the country from the function call
    CNTRY = CNTRY,
    # the item to download
    item = item,
    # the username from this key's record
    username = keyring::key_list("human-fertility-database")$username,
    # the password for this key's record
    password = keyring::key_get(
      service = "human-fertility-database",
      username = keyring::key_list("human-fertility-database")$username
    )
  )
}

# Download a data set iteratively for all named countries using purrr::map()
read_hfd_countries_item <- function(countries, item){
    countries %>%
        # Returns a list of data.frames, adding a column for country code to each
        # the map() function performs a run of Ben's read_hmd_country() 
        #   function for each listed country
        purrr::map_dfr(function(ccode) {
            # read item
            read_hfd_country(ccode, item) %>%
                # this adds the column "country" storing the country ISO code
                dplyr::mutate(ccode = ccode)
        }) %>%
        # Phil added this to make it a tibble
        tibble() %>% 
        # and add country name
        left_join(hfdcodes, by = "ccode")
}
```

Pull CBR and TFR data for United States, Lithuania, and Japan:

```{r}
# look at US, Lithuania, Japan
countries <- hfdcodes %>% 
    filter(Name %in% c("United States", "Lithuania", "Japan")) %>% 
    pull(ccode)

# call crude birth rates
cbr <- read_hfd_countries_item(countries = countries, item = "cbrRR")

# call total fertility rates
tfr <- read_hfd_countries_item(countries = countries, item = "tfrRR")

# combine datasets
df <- cbr %>% left_join(tfr, by = c("Year","Name","ccode"))
```

Plot CBR and TFR over time in each country:

```{r}
# plot CBR over time
cbr_plot <- qplot(Year,CBR,data=df,color = Name,geom="line")

# plot TFR over time
tfr_plot <- qplot(Year,TFR,data=df,color = Name,geom="line")

# arrange plots in a nice layout
ggarrange(cbr_plot, tfr_plot,ncol = 2, nrow = 1,common.legend = TRUE, legend="bottom")
```
\
_`r figure_nums(name = "cbr-tfr", caption = "Yearly CBR and TFR in 3 countries.")`_

From  `r figure_nums(name="cbr-tfr",display="cite")` we see a general trend of declining crude birth rates and total fertility rates. However, these two measures are not exact proxies for each other. Total fertility rates do not depend on age-structure; they are calculated as a sum of age-specific fertility rates. On the other hand, crude birth rates depend on population structure: populations with more women of childbearing age will experience higher birth rates when controlling for total fertility rates. To get a sense of the different information encoded in these rates, we now plot the ratio of CBR and TFR over time in each country:

```{r}
# plot ratio CBR/TFR over time
ratio <- qplot(Year,CBR/TFR,data=df,color = Name,geom="line")
ggarrange(ratio,ncol = 1, nrow = 1,common.legend = TRUE, legend="bottom")
```
\
_`r figure_nums(name = "ratio", caption = "Ratio CBR/TFR in 3 countries.")`_





# Source code

```{r, ref.label=knitr::all_labels(),echo=TRUE,eval=FALSE}
```

